Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-07-09 thirdPaper/draft/typeConeAssociahedra.tex
began main results

@@

\subsection{Polytopes and fans}

@@

\subsection{Type cone}

>>>>

\section*{Main results and logical dependencies}

We give a brief overview of our main results.
Their logical dependencies are recalled in a schematic diagram at the end of this section.

Let~$\Fan$ be an essential, complete, simplicial, polyhedral fan in~$\R^n$ (all definitions are recalled in \cref{ssec:polytopes and fans}).
The type cone~$\typeCone(\Fan)$ of~$\Fan$ (\cref{ssec:type cone}), introduced by P. McMullen, is a device that encodes the information about all possible polytopal realizations of~$\Fan$.
If the simplicial fan~$\Fan$ has the unique exchange relation property (\cref{def:uerp}), then its type cone is easier to describe and study.
In particular, if moreover the type cone~$\typeCone(\Fan)$ is simplicial (\cref{rem:dimTypeCone}), then all possible polytopal realizations of~$\Fan$ can be described explicitely in~$\R^N_{>0}$, where~$N$ is the number of rays of~$\typeCone(\Fan)$, in terms of~$N-n$ positive real parameters (\cref{coro:simplicialTypeCone}).
This gives a new point of view on~\cite{ArkaniHamedBaiHeYan,BazierMatteDouvilleMousavandThomasYildirim}, illustrated in \cref{subsec:typeConeAsso} for classical associahedra, that we then apply to two families of~$\b{g}$-vector fans generalizing the~$\b{g}$-vector fans of classical associahedra: The~$\b{g}$-vector fans of cluster algebras of finite type (generalized associahedra), and the~$\b{g}$-vector fans of gentle algebras (non-kissing associahedra, non-crossing associahedra, generalized accordiohedra).
We also describe the type cone for graph associahedra (\cref{subsec:typeConeGA}), to which our strategy does not apply.

\subsection*{Cluster algebras of finite type} (\cref{subsec:typeConeCA})

We first consider the~$\b{g}$-vector fan~$\Fan(\B_\circ)$ of a skew-symmetrizable cluster algebra of finite type, with respect to any initial exchange matrix of finite type~$\B_\circ$.
A reformulation of~\cite[Thm 7.5]{BuanMarshReinekeReitenTodorov}
\medskip

\noindent \ref{prop:uniqueExchangePropertyCA}: The cluster fan~$\Fan(\B_\circ)$ has the unique exchange relation property.

\medskip

We consider the set~$\meshes$ of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and we denote by~$\principalVariables$ the set of cluster variables.
Let~$\{x,x'\}\in\meshes$, and let~$X,X'$ be two clusters such that~$X\setminus\{x\}=X'\setminus\{x'\}$.
For any~$y\in\principalVariables$, define~$\alpha_{x,x',y}$ to be~$|b_{xy}|$ if~$y\in X\cap X'$ and 0 otherwise.
By the unique exchange relation property, the~$\alpha_{x,x',y}$'s do not depend on the choice of the clusters~$X,X'$.

\medskip

\noindent \cref{thm:allPolytopalRealizationsCA}: 
\emph{For any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \alpha_{x,x',y} \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.}

\subsection*{Gentle algebras}

\subsection*{Cluster categories}

\subsection*{Extriangulated categories}

@@

\subsection{Polytopes and fans}
\label{ssec:polytopes and fans}

@@

\subsection{Type cone}
\label{ssec:type cone}
